A fast unified model for parsing and sentence understanding

### 2016-03-15 writing/hard_stack_paper/paper.tex
Initial revision of abstract and Section 1.

Tree-structured neural networks exploit valuable syntactic parse information as they interpret the meanings of sentences. However, they suffer from two key technical problems that make them slow and unwieldy for large-scale NLP tasks: they can only operate on parsed sentences and do not directly support batched computation. We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single model by integrating tree-structured sentence interpretation into the linear sequential structure of a shift-reduce parser. This model  supports batched computation, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy. We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models. \todo{[SB] Quantify success. Emphasize novelty.}

\end{abstract}

@@

A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \citep{socher2011semi,sutskever2014sequence}. This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vector to a sentence vector, and this function can take a range of different forms. Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory (LSTM RNNs, \citet{hochreiter1997long}, see Figure~\ref{fig:batching:good}), which accumulate information over the sentence sequentially, convolutional neural networks over words or characters (\citet{kalchbrenner2014convolutional}, \todo{[SB] Cite LeCun CharCNN}), which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks (TreeRNNs \citet{goller1996learning,socher2011semi}, see Figure~\ref{fig:batching:bad}), which propagate information up a binary parse tree structure over words. Of these, it is easy to argue that the TreeRNN is the principled choice---meaning in natural language sentences is widely believed to be constructed incrementally according to a tree structure \citep{Partee84,Janssen97}, and so a model which instantiates that tree structure should have a maximally easy time learning to construct meaning. \todo{[SB] Trim this para.}

TreeRNNs have shown promise, but they have largely been overlooked in favor of sequence-based RNNs for practical reasons. Batched computation---performing synchronized computation across many examples at once---is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets (\question{What, if anything, can we cite here?} \todo{[SB] Bengio 2003?}). Batching yields order-of-magnitude improvements in standard CPU implementations of neural network models, and substantial additional gains on GPUs. TreeRNNs build the parse structure of each sentence they process into the structure of the network itself, forcing the model to perform a different sequence of computations for each sentence, and thereby requiring that each sentence be processed one-by-one, making batched computation impossible in standard implementations. One aim of this paper is to restructure the computations involved in a TreeRNN in order to enable batching and make these large-scale evaluations possible.  In addition to their slow speed, TreeRNNs have an additional feature that makes them impractical in many settings: they can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the implementation of these models for most applications. This paper additionally aims to integrate parsing into TreeRNN processing to circumvent this issue. \todo{[SB] Trim this para. Reframe parsing claim around Socher 2011.}

In spite of the unwieldiness of current standard TreeRNNs, there is still value in the idea of using syntactic parse trees to guide the construction of meaning in neural network models. Tree-structured models have shown substantial performance gains on some smaller-scale NLP tasks \citep{tai2015improved,li2015tree}, and their theoretical foundations are strong. Tree-structured compositionality is a property of natural language, and simpler sequence-based RNNs have a difficult time exploiting this property---while sequence-based models should, theoretically, be able to learn to simulate tree-structured composition, this behavior does not emerge in practice on reasonably-sized tasks \citep{bowman2015trees}.

In this paper, we introduce the Stack-augmented Parser-Interpreter Neural Network, or SPINN. The SPINN executes tree-structured computations in a linearized sequence, and can incorporate a neural network parser that determines the overall tree structure on-the-fly. This design improves upon the TreeRNN architecture in three ways:
\begin{itemize}
\item At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser.
\item It supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs.
\item It supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.
\end{itemize}

\todo{[SB] Move the introduction of the SPINN earlier in the introduction. Mention that we'll be using SNLI as our sole evaluation task.}

\section{The shift-reduce sentence model}

@@

\subsection{Related work: Transition-based parsing and neural networks.}
\todo{[SB] Move!}

There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing, of which shift-reduce parsing is a version \citep{henderson2004discriminative,emami2005neural,titov2010latent,buys2generative,chen2014,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy} \todo{[SB] Order citations by year}. In addition, there has been recent work \citep{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well. To our knowledge, the SPINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.

>>>>

Tree-structured neural networks exploit valuable syntactic parse information as they interpret the meanings of sentences. However, they suffer from two key technical problems that make them slow and unwieldy for large-scale NLP tasks: they can only operate on parsed sentences and do not directly support batched computation. We address these issues by introducing the Stack-augmented Parser-Interpreter Neural Network (SPINN), which combines parsing and interpretation within a single tree-sequence hybrid model by integrating tree-structured sentence interpretation into the linear sequential structure of a shift-reduce parser. Our model supports batched computation for a speedup of up to 20x over other tree-structured models, and its integrated parser allows it to operate on unparsed data with nearly no loss of accuracy. We evaluate it on the Stanford NLI entailment task and show that it significantly outperforms other sentence encoding models.
\end{abstract}

@@

A wide range of current state-of-the-art models in NLP are built around neural network components that build vector representations of sentence meaning \citep{socher2011semi,sutskever2014sequence}. This component, the sentence encoder, is generally formulated as a learned parametric function from a sequence of word vector to a sentence vector, and this function can take a range of different forms. Common sentence encoders include sequence-based recurrent neural network models (RNNs) with Long Short-Term Memory \citep[LSTM,][see Figure~\ref{fig:batching:good}]{hochreiter1997long}, which accumulate information over the sentence sequentially, convolutional neural networks over words or characters \citep{kalchbrenner2014convolutional,DBLP:journals/corr/ZhangZL15}, which accumulate information using filters over short local sequences of words, and tree-structured recursive neural networks \citep[TreeRNNs,][see Figure~\ref{fig:batching:bad}]{goller1996learning,socher2011semi}, which propagate information up a binary parse tree structure over words.

Of these, it is easy to argue that the TreeRNN is the principled choice, since meaning in natural language sentences constructed incrementally according to a tree structure \citep{Partee84,Janssen97}. However, while these models have shown promise \citep{tai2015improved,li2015tree,bowman2015trees}, they have largely been overlooked in favor of sequence-based RNNs because of their incompatibility with batched computation and their reliance on external parsers.  Batched computation---performing synchronized computation across many examples at once---yields order-of-magnitude improvements in model run time, and is a crucial enabling technology in allowing neural networks to be trained efficiently on large datasets. Because TreeRNNs use a different model structure for each sentence, batched computation is impossible in standard implementations. In addition, standard TreeRNN models can only operate on sentences that have already been processed by a syntactic parser, which slows and complicates the implementation of these models for most applications.

This paper introduces a new model to address both these issues: the Stack-augmented Parser-Interpreter Neural Network, or SPINN, shown in Figure~\ref{fig:m1-views}. The SPINN executes the computations of a tree-structured model in a linearized sequence of operations, and can incorporate a neural network parser that determines the overall tree structure on-the-fly. This design improves upon the TreeRNN architecture in three ways:
\begin{itemize}
\item At test time, it can simultaneously parse and interpret unparsed sentences, removing the dependence on an external parser.
\item It supports batched computation for both parsed and unparsed sentences, supporting dramatic speedups over standard TreeRNNs.
\item It supports a novel tree-sequence hybrid mechanism for handling local context in sentence interpretation that yields substantial gains in accuracy over either pure sequence-based models or pure tree-based models.
\end{itemize}

We evaluate the SPINN on the Stanford Natural Language Inference sentence understanding task \citep[SNLI,][]{snli:emnlp2015}, and find that it significantly outperforms other sentence encoding-based models, and that it yields speed increases of upwards of 10x over comparable TreeRNN models.

\subsection{Related work}

There has been a fairly long history of work on building neural-network based parsers that use the core operations and data structures from transition-based parsing, of which shift-reduce parsing is a version \citep{henderson2004discriminative,emami2005neural,titov2010latent,chen2014,buys2generative,dyer-EtAl:2015:ACL-IJCNLP,kiperwasser2016easy}. In addition, there has been recent work \citep{zhang2016top,dyer2016rnn} proposing models designed primarily for generative language modeling tasks that use these structures as well. To our knowledge, the SPINN is the first model to use these structures for the purpose of sentence interpretation, rather than parsing or generation.

\citet{socher2011dynamic} presents a version of the TreeRNN model which is capable of operating over unparsed inputs. However, this method is unsupervised, limiting its performance, and it requires an expensive search process at test time. Our model presents an alternative approach to combining interpretation and parsing that supports supervised training for the parsing component, and that is dramatically faster than any previous tree-structured model.

\section{Our model: The SPINN}

@@
